---
title: "P8105 Homework 3"
author: "Alice Tivarovsky"
date: "10/6/2019"
output: github_document
---

## Setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Problem 1

### Instacart Data

Loading data and generating preview: 
```{r}
library(p8105.datasets)
data("instacart")
view(head(instacart, n = 50))
```


The instacart dataset contains 1,384,617 observations and 15 variables. It catalogues instacart orders from October 2017 in New York City. Key variables include order_id (the order identifier), user_id (the customer identifier), product_id (the product name within each order), and order_dow (the day of the week). The first row, for example, is order_id = 1, meaning it was the first order in October 2017, includes (among 7 other items) the item product_name = Bulgarian Yogurt, which was ordered by customer user_id = 112108 on order_dow = 4 (presumably Wednesday). 

## EDA with Instacart Dataset

Counting # of aisles and aisles most commonly ordered from. 
```{r}
pop_aisle = 
  count(instacart, aisle_id) %>% 
  arrange(desc(n))

pop_aisle
```

There are 134 aisles and the aisle most commonly ordered from is aisle 83 (150609 items). 

Keeping only aisles with n > 10,000, creating a plot that shows the number of items ordered in each aisle. 

```{r}

```










